%============================================================================
% nomenclature.tex
%============================================================================
% This file contains:
% 	- List of defined nomenclature items

\printnomenclature[6em]

\nomtypeA{NRC}{Nuclear Regulatory Commision}
\nomtypeA{NPP}{Nuclear Power Plant}
\nomtypeA{10CFR}{Title 10 of the Code of Federal Regulations}
\nomtypeA{10CFR50}{Part 50 of 10CFR}
\nomtypeA{SAR}{Safety Analysis Report}
\nomtypeA{LWR}{Light Water Reactor}
\nomtypeA{ECCS}{Emergency Core Cooling System}
\nomtypeA{LOCA}{Loss-of-Coolant Accident}
\nomtypeA{PWR}{Pressurized Water Reactor}
\nomtypeA{BWR}{Boiling Water Reactors}
\nomtypeA{CFL}{Courant-Friedrichs-Lewy}
\nomtypeA{NBC}{Nonlinear-Boundary Continuity}
\nomtypeA{PDE}{Partial Differential Equation}
\nomtypeA{SETS}{Stability Enhancing Two-Step}
\nomtypeA{PVM}{Parallel Virtual Machine}

\nomtypeR{$w^{n+1}_{k, j}$}{Volumetric flow for phase k in \relap53d{}.}{$ \text{ft}^{3}\,\text{s}^{-1}$}
\nomtypeR{$u^{n+1}_{k, j}$}{Flow of internal energy for phase k in \relap53d{}.}{$\text{BTU}\,\text{s}^{-1} $}
\nomtypeR{$n^{n+1}_{g,j}$}{Flow of \ncg{} mass in \relap53d{}.}{$ \lbm{} \, \text{s}^{-1}$}
\nomtypeR{$m^{n+1}_{k,j}$}{Flow of mass for phase k in \relap53d{}.}{$ \lbm{}\, \text{s}^{-1} $}
\nomtypeR{$s_{p}$}{Source or sink of external momentum.}{$ \lbf{} \text{s}$}
\nomtypeR{$s_{e}$}{Source or sink of external energy.}{$ \text{BTU} $}
\nomtypeR{$s_{m}$}{Source or sink of external mass.}{$\lbm{} $}
\nomtypeR{$\vec{a}$}{Vector of constants from right-hand-side of momentum system.}{}
\nomtypeR{$\vec{b}$}{Vector of pressure-update coefficients from momentum system.}{}
\nomtypeR{$\vec{u}$}{Vector of phasic velocities for a given flow path.}{$ \text{ft}\, \text{s}^{-1}$}
\nomtypeR{$\vec{x}$}{Vector of nine independent variables used by \cobra{}.}{}
\nomtypeR{$\momVec{}$}{Vector of three momenta fields.}{$ \lbm{}\, \text{s}^{-1}$}
\nomtypeR{$\vec{E}^{*}$}{Approximation of temporal integral of $\vec{E}(\vec{y}(\vec{x}))$.}{}
\nomtypeR{$\vec{S}$}{Vector of operator based scale factors.}{}
\nomtypeR{$\vec{F}$}{Vector of nonlinear residuals.}{}
\nomtypeR{$\vec{E}$}{Vector of spatially discrete conservation equations.}{}
\nomtypeR{$\vec{e}$}{Vector of conservation equations, except the temporal derivatives.}{}
\nomtypeR{$\vec{y}$}{Vector of conserved quantities.}{}
\nomtypeR{$\vec{g}$}{Gravitational acceleration vector.}{$ \text{ft} \, \text{s}^{-2} $}
\nomtypeR{$\vec{J}$}{A given Jacobian matrix.}{}
\nomtypeR{$\vec{I}$}{Identity matrix.}{}
\nomtypeR{$\vec{K}$}{Matrix representing inter-continuity volume pressure coupling coefficients.}{}
\nomtypeR{$\vec{C}$}{Vector of unknowns for a given continuity volume.}{}
\nomtypeR{$\vec{r}$}{Vector of right-hand sides for a given continuity volume.}{}
\nomtypeR{$\vec{U}$}{Upper triangular matrix from \vec{LU} decomposition.}{}
\nomtypeR{$\vec{L}$}{Lower triangular matrix from \vec{LU} decomposition.}{}
\nomtypeR{$\vec{A}$}{Pressure matrix.}{}
\nomtypeR{$\vec{res}$}{Right hand side of pressure system.}{}
\nomtypeR{$\vec{Q}$}{Matrix representing inter-domain flow rate coefficients.}{}
\nomtypeR{$\vec{B}$}{Matrix representing pressure matrix inter-domain coupling coefficients.}{}
\nomtypeR{$\vec{Z}$}{Matrix representing a linear domain continuity volume's linear system.}{}
\nomtypeR{$\vec{W}$}{Matrix representing global inter-domain coupling coefficients.}{}
\nomtypeR{$\vec{w}$}{Vector representing inter-domain coupling coefficients for a given NBC volume.}{}
\nomtypeR{$V$}{Volume of a given continuity volume.}{$ \text{ft}^{3} $}
\nomtypeR{$x$}{Spatial location.}{$\text{ft}$}
\nomtypeR[A]{$A$}{Cross-sectional area.}{$ \text{ft}^{2} $}
\nomtypeR[A]{$\tilde{A}$}{Effective cross-sectional area between two flow paths.}{$ \text{ft}^{2} $}
\nomtypeR[A]{$A(t)_{r}$}{Time dependent area ratio in valve problem.}{}
\nomtypeR{$K(t)$}{Effective time dependent loss coefficient in valve problem.}{$\lbf{}\,\text{s}\,\text{ft}^{-1}$}
\nomtypeR{$v$}{Specific volume.}{$\text{ft}^{3} \, \lbm{}$}
\nomtypeR{$K_{o}$}{Base loss coefficient in valve problem.}{$\lbf{}\,\text{s}\,\text{ft}^{-1}$}
\nomtypeR{$T_{\text{CPU}}$}{CPU time taken for a simulation to run.}{$\text{s}$}
\nomtypeR{$R$}{Integral of nonlinear residuals.}{}
\nomtypeR{$\tilde{R}$}{Time-averaged integral of nonlinear residuals.}{}
\nomtypeR{$\tilde{R}_{M}$}{Time-moment integral of nonlinear residuals.}{}
\nomtypeR{$q_{n,l}$}{Energy transferred between the \ncg{} field and the liquid water phase.}{$ \text{BTU}$}
\nomtypeR{$q_{w,\phi}$}{Energy transferred between a solid structure and a fluid $\phi$.}{$\text{BTU}$}
\nomtypeR{$q_{i,\phi}$}{Energy transferred between the saturated interface and the fluid $\phi$.}{$ \text{BTU}$}
\nomtypeR{$c$}{Local speed of sound.}{$ \text{ft}\, \text{s}^{-1} $}
\nomtypeR{$\mathcal{L}_{\infty}$}{Infinity norm of a given vector.}{}
\nomtypeR{$\mathcal{L}_{2}$}{Euclidean norm of a given vector.}{}
\nomtypeR{$N_{\text{CPL}}$}{Number of \cobra{} -- \relap53d{} coupling interfaces.}{}
\nomtypeR{$N_{f}$}{Set of flow paths connected to a given continuity volume.}{}
\nomtypeR{$N_{s}$}{Number of stages in a multi-stage temporal integration scheme.}{}
\nomtypeR{$N_{t}$}{Number of successful timesteps in a simulated time span.}{}
\nomtypeR{$n_{t}$}{Number of time step refinements in timestep generation function.}{}
\nomtypeR{$N_{n}$}{Number of inter-domain connections in the global domain.}{}
\nomtypeR{$N_{u}$}{Number of variables in a residual or update vector.}{}
\nomtypeR{$N_{\text{lin}}$}{Number of continuity volumes in the linear domain.}{}
\nomtypeR{$N_{\text{nln}}$}{Number of continuity volumes in the nonlinear domain.}{}
\nomtypeR{$N_{\text{NBC}}$}{Set of inter-domain connection for a linear continuity volumes.}{}
\nomtypeR{$N_{c}$}{Set of continuity volumes connected to a given flow path.}{}
\nomtypeR{$S$}{Scale factor for an equation.}{}
\nomtypeR{$t$}{Time.}{$\text{s}$}
\nomtypeR{$C_{1}$}{Coefficient in phase transition function.}{}
\nomtypeR{$P$}{Pressure.}{$\text{psia}$}
\nomtypeR{$h$}{Specific enthalpy.}{$ \text{BTU}\, \lbm{}^{-1}$}
\nomtypeR{$f(x)$}{A generic function.}{}
\nomtypeR{$r_{f}$}{Refinement factor for timestep generation function.}{}
\nomtypeR{$\dot{m}$}{A generic momentum.}{$\lbf{}$}
\nomtypeR{$u$}{A generic velocity.}{$\text{ft}\, \text{s}^{-1}$}
\nomtypeR{$K_{i,\phi_1 \phi_2}$}{Effective coefficient for drag between fluids $\phi_1$ and $\phi_2$.}{$\lbf{}\,\text{s}\,\text{ft}^{-1}$}
\nomtypeR{$K_{w,\phi}$}{Effective coefficient for drag between wall and fluid $\phi$.}{$\lbf{}\,\text{s}\,\text{ft}^{-1}$}
\nomtypeR{$e_{p}$}{Error in calculated pressure}{$ \text{psia} $}
\nomtypeR{$T$}{End time of simulation.}{$\text{s} $}

\nomtypeG{$F_{\text{tol}}$}{Convergence tolerance with respect to the scaled residual vector.}{}
\nomtypeG{$\dtmax{}$}{Maximum timestep allowed for a simulation.}{$\text{s}$}
\nomtypeG{$\dt{}$}{Timestep between time $t^{n}$ and $t^{n+1}$.}{$\text{s}$}
\nomtypeG{$\dx{}$}{Axial length of a volume.}{$\text{ft}$}
\nomtypeG{$\eta$}{Apportionment factor for inter-phase mass transfer.}{}
\nomtypeG{$\Upsilon$}{Inter-field field source or sink of mass.}{$\lbm{}$}
\nomtypeG{$\rho$}{Density.}{$\lbm{} \, \text{ft}^{-3} $}
\nomtypeG{$\Gamma$}{Inter-phase source or sink of mass.}{$ \lbm{}$}
\nomtypeG{$\alpha$}{Volume fraction.}{}
\nomtypeG{$\delta \vec{x}$}{Vector of updates to the independent parameters.}{}
\nomtypeG{$\tau^{'}_{w,\phi}$}{Shear force from contact between the channel wall and fluid $\phi$.}{$\lbf{} \, \text{ft}^{-3}$}
\nomtypeG{$\tau^{'}_{i,\phi_{1} \phi_{2}}$}{Shear force from contact between the field or phase $\phi_{1}$ and $\phi_{2}$.}{$ \lbf{} \, \text{ft}^{-3} $}
\nomtypeG{$\Gamma \vec{u}^{'}$}{Momentum transfer due to the exchange of mass between the aqueous phases.}{$ \lbf{} $}
\nomtypeG{$\Upsilon \vec{u}^{'}$}{Momentum transfer due to the exchange of mass between the two liquid fields.}{$ \lbf{} $}
\nomtypeG{$\Gamma h^{'}_{\phi}$}{Energy transfer rate due to aqueous phase change.}{$ \text{BTU}$}
\nomtypeG{$\vec{\Xi}$}{Matrix of coefficients that converts the momentum vector to the flow rate vector.}{}
\nomtypeG{$\Psi$}{A given flow rate.}{}
\nomtypeG{$\vec{\Psi}$}{Vector of flow rates.}{}
\nomtypeG{$\Omega$}{The domain; the set of volumes that determines the domain.}{}
\nomtypeG{$\delta_{\text{tol}}$}{Convergence tolerance with respect to the relative update vector.}{}

\nomtypeT{k}{Newton iteration index.}
\nomtypeT{n}{Time index.}
\nomtypeT{s}{Temporal integral stage index.}
\nomtypeT{$'''$}{Per unit volume.}
\nomtypeT{b}{The time point at which the donored quantities are evaluated in the donoring operator.}
\nomtypeT{c}{The time point at which the donoring velocity is evaluated in the donoring operator.}
\nomtypeT{.}{Per second, a denotation of a rate.}
\nomtypeT{$C_{2}$}{Coefficient in phase transition function.}
\nomtypeT{p}{Generic power.}

\nomtypeS{$\phi$}{A generic field or phase depending upon context.}
\nomtypeS{n}{Indicating the \ncg{} component of the gaseous phase.}
\nomtypeS{g}{Indicating the gaseous phase.}
\nomtypeS{e}{Indicating the entrained liquid field.}
\nomtypeS{l}{Indicating either the continuous liquid field or the total liquid phase.}
\nomtypeS{m}{Relating to a momentum flow path.}
\nomtypeS{v}{Indicating the vapor component of the gaseous phase.}
\nomtypeS{c}{Relating to a continuity volume.}
\nomtypeS{e}{Relating to a continuity volume's energy equations.}
\nomtypeS{d}{Donored quantity.}
\nomtypeS{lin}{Relating to the linear domain or solver.}
\nomtypeS{nln}{Relating to the nonlinear domain or solver.}
\nomtypeS{a}{Averaged quantity.}
\nomtypeS{r}{Relative.}
\nomtypeS{dep}{Depleting phase.}
\nomtypeS{car}{Carrier phase.}
\nomtypeS{p}{Index representing a particular coordinate for a given inter-domain flow path.}
\nomtypeS{j}{The coordinate of a particular continuity volume.}
\nomtypeS{i}{Generic index representing a particular coordinate of a location.}
\nomtypeS{o(i)}{Coordinate of continuity volume connected via a flow path at coordinate i.}
\nomtypeS{s(i)}{Coordinate of continuity volume connected via a flow path at coordinate i.}
